Measuring causal influence with\\ back-to-back regression: the linear case

### 2019-11-13 paper/paper.tex
small change

% $$
\begin{equation}
  X^\top X = \left(\begin{array}{lccl}\Sigma_{X_1 X_1} & \Sigma_{X_1 X_2} \\ \Sigma_{X_1 X_2} & \Sigma_{X_2 X_2}\end{array}\right),\qquad\qquad N^\top N = \left(\begin{array}{lccl}\Sigma_{N_1 N_1} & \Sigma_{N_1 N_2} \\ \Sigma_{N_1 N_2} & \Sigma_{N_2 N_2}\end{array}\right),
\end{equation}
% $$
% and
\begin{equation}
  \hat H = \left(\begin{array}{cc}(\Sigma_{X_1 X_1}+\Sigma_{N_1 N_1})^{-1}\Sigma_{X_1 X_1} & (\Sigma_{X_1 X_1}+\Sigma_{N_1 N_1})^{-1}\Sigma_{X_1 X_2} \\0 & 0\end{array}\right),
\end{equation}
\begin{equation}
  \text{diag}_k (\hat H) = \text{diag}((\Sigma_{X_1 X_1}+\Sigma_{N_1 N_1})^{-1}\Sigma_{X_1 X_1}) = \text{diag}((I+\Sigma_{X_1 X_1}^{-1}\Sigma_{N_1 N_1})^{-1}).
\end{equation}
%
% $$
% $$
% $$
% $$
%
In the absence of noise, we have $\Sigma_{N_1 N_1}=0$, and so $\text{diag}_k(\hat H)=I$, and $$\text{diag}(\hat H) = \text{diag}(E)$$
Otherwise the $k$ first diagonal elements of $\hat H$ are all positive (since we assume that $F$ has full rank over $Img(E)$, they cannot be zero), lower than $1$, and bounded by
$$\frac{\sigma_{X_k}}{\sigma_{X_k} +\sigma_{N_1}} \text{ and }  \frac{\sigma_{X_1}}{\sigma_{X_1} +\sigma_{N_k}}$$where  $\sigma_{X_1}$, $\sigma_{X_k}$, $\sigma_{N_1}$ and $\sigma_{N_k}$ denote the largest and smallest eigenvalues of $\Sigma_{X_1 X_1}$ and $\Sigma_{N_1 N_1}$.

In practical cases, non linearities and sampling error will add noise to the diagonal elements of $\hat H$. To test diagonal elements for equality with zero, we note that the average value of the non-zero coefficients of $\text{diag}(\hat H)$ is the trace of $\hat H$ divided by $k$. Since $X$ and $N$ are not correlated, $\Sigma_{X_1 X_1}^{-1}\Sigma_{N_1 N_1}$ should have a low condition number, which means $\frac{Var(X)}{Var(X)+Var(N)}$ should provide a good estimate of $\frac{1}{k}Tr(\hat H)$. This can be used to test whether a feature of $X$ has causal influence. An alternative is the Sonquist-Morgan criterion presented in the appendix.

>>>>

% $$
\begin{equation}
  X^\top X = \left(\begin{array}{lccl}\Sigma_{X_1 X_1} & \Sigma_{X_1 X_2} \\ \Sigma_{X_1 X_2} & \Sigma_{X_2 X_2}\end{array}\right),\qquad\qquad N^\top N = \left(\begin{array}{lccl}\Sigma_{N_1 N_1} & \Sigma_{N_1 N_2} \\ \Sigma_{N_1 N_2} & \Sigma_{N_2 N_2}\end{array}\right),
\end{equation}
% $$
% and
\begin{equation}
  \hat H = \left(\begin{array}{cc}(\Sigma_{X_1 X_1}+\Sigma_{N_1 N_1})^{-1}\Sigma_{X_1 X_1} & (\Sigma_{X_1 X_1}+\Sigma_{N_1 N_1})^{-1}\Sigma_{X_1 X_2} \\0 & 0\end{array}\right),
\end{equation}
\begin{equation}
  \text{diag}_k (\hat H) = \text{diag}((\Sigma_{X_1 X_1}+\Sigma_{N_1 N_1})^{-1}\Sigma_{X_1 X_1}) = \text{diag}((I+\Sigma_{X_1 X_1}^{-1}\Sigma_{N_1 N_1})^{-1}).
\end{equation}
%
% $$
% $$
% $$
% $$
%
In the absence of noise, we have $\Sigma_{N_1 N_1}=0$, and so $\text{diag}_k(\hat H)=I$, and $$\text{diag}(\hat H) = \text{diag}(E)$$
Otherwise the $k$ first diagonal elements of $\hat H$ are all positive, since we assume that $F$ has full rank over $Img(E)$. In addition, they are bounded:
$$\frac{\sigma_{X_k}}{\sigma_{X_k} +\sigma_{N_1}} \leq \text{diag}_k(\hat{H})\leq  \frac{\sigma_{X_1}}{\sigma_{X_1} +\sigma_{N_k}}$$
where  $\sigma_{X_1}$, $\sigma_{X_k}$, $\sigma_{N_1}$ and $\sigma_{N_k}$ denote the largest and smallest eigenvalues of $\Sigma_{X_1 X_1}$ and $\Sigma_{N_1 N_1}$. This allows us to recover $E$ by identifying the non-zero elements on the diagonal of $\hat{H}$.

In practical cases, non-linearities and sampling error will add noise to the diagonal elements of $\hat H$. We need to test whether diagonal elements are different from zero. The average value $\mu$ of non-zero coefficients of $\text{diag}(\hat H)$ is the trace of $\hat H$ divided by $k$. $\mu$ can be estimated as follow :
\begin{equation}
$\mu = \frac{Var(X)}{Var(X)+Var(N)}$
\end{equation}
This provides a method to identify causal features
%Since $X$ and $N$ are not correlated, $\Sigma_{X_1 X_1}^{-1}\Sigma_{N_1 N_1}$ should have a low condition number, which means  should provide a good estimate of $\frac{1}{k}Tr(\hat H)$.

To test whether a feature of $X$ has causal influence. An alternative is the Sonquist-Morgan criterion presented in the appendix.
